\RequirePackage{snapshot}
%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a contributed volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[graybox]{svmult}

% choose options for [] as required from the list
% in the Reference Guide

\usepackage{mathptmx}       % selects Times Roman as basic font
\usepackage{helvet}         % selects Helvetica as sans-serif font
\usepackage{courier}        % selects Courier as typewriter font
\usepackage{type1cm}        % activate if the above 3 fonts are
                            % not available on your system
%
\usepackage{makeidx}         % allows index generation
\usepackage{graphicx}        % standard LaTeX graphics tool
                             % when including figure files
\usepackage{multicol}        % used for the two-column index
\usepackage[bottom]{footmisc}% places footnotes at page bottom

% see the list of further useful packages
% in the Reference Guide

\makeindex             % used for the subject index
                       % please use the style svind.ist with
                       % your makeindex program



\input{packages-springer.tex}
\input{macro-springer.tex}

\graphicspath{{./figure/}}

% only print used references
% WARNING: does not work with showlabels
%\mathtoolsset{showonlyrefs=true}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\title*{On solving frictional contact problems: formulations and comparisons of numerical methods.}
\titlerunning{On solving frictional contact problems.}
% Use \titlerunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\author{Vincent Acary and Maurice Br\'emond and Olivier Huber}
% Use \authorrunning{Short Title} for an abbreviated version of
% your contribution title if the original one is too long
\institute{Vincent Acary \at INRIA, 655, Avenue de l'Europe, 38334 Saint Ismier Cedex, France \email{vincent.acary@inria.fr}
  \and Maurice Br\'emond \at INRIA, 655, Avenue de l'Europe, 38334 Saint Ismier Cedex, France \email{maurice.bremond@inria.fr}
\and Olivier Huber  \at Wisconsin Institute for Discovery, University of Wisconsin-Madison, 330 N. Orchard St., 53715, USA \email{ohuber2@wisc.edu}}
%
% Use the package "url.sty" to avoid
% problems with special characters
% used in your e-mail or web address
%
\maketitle


\def\ourabstract{
In this chapter, we review several formulations of the discrete frictional contact problem that arises in space and time discretized mechanical systems with unilateral contact and three-dimensional Coulomb's friction. Most of these formulations are well--known concepts in the optimization community, or more generally, in the mathematical programming community. To cite a few, the discrete frictional contact problem is formulated as variational inequalities, generalized or semi--smooth equations, second--order cone complementarity problems, or as optimization problems such as quadratic programming problems over second-order cones. Thanks to these multiple formulations, various numerical methods emerge naturally for solving the problem. We review the main numerical techniques that are well-known in the literature and we also propose new applications of methods such as the fixed point and extra-gradient methods with self-adaptive step rules for variational inequalities or the proximal point algorithm for generalized equations. All these numerical techniques are compared over a large set of test examples using performance profiles. One of the main conclusion is that there is no universal solver. Nevertheless, we are able to give some hints to choose a solver with respect to the main characteristics of the set of tests.}


\abstract*{\ourabstract}

\abstract{\ourabstract}


\input{body.tex}


\begin{acknowledgement}
The authors are grateful to Florent Cadoux, Fr\'ed\'eric Dubois, Claude L\'emar\'echal, J\'er\^ome Malick and Mathieu Renouf  for stimulating discussions.
\end{acknowledgement}
%

\input{appendix.tex}

\bibliographystyle{spbasic}
%\bibliography{./biblio/String,./biblio/NonSmooth,./biblio/Math,./biblio/Multibody,./biblio/Fem.bib,./biblio/Dae.bib,./biblio/Meca,./biblio/AnaNum.bib,./biblio/Math-Impact,./biblio/Contact,./biblio/Optim,./biblio/Cp}
\bibliography{biblio-extract}


%\input{referenc}
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
